### 2022-09-02 main/chapter_searches.tex
add some words on frequencies

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of discrete randomly occuring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$, and
$N$ have satisfied $b$.
From Equation~\ref{eqn:searches_binomial}, this converges to the general
probability in its infinite limit.

@@

\TODO{
on ``likelihood'' models ---
we are instructed to pretend that these are likelihoods on auxiliary data
TODO cite.
One is free to engage in such doublethink, just as one is free to proclaim that
$2 + 2 = 5$.
The fact remains that many constraints do not correspond to predictions of
auxiliary data: counterexamples include
scale variations,
theoretical cross-section uncertainties,
non-closure uncertainties (which cover differences between estimates).
simulated sample camparisons,
statistical errors on weighted simulations,
extrapolation uncertainties,
and
all uncertainties in supersymmetric signal yields
(for which we have yet no data).
}

>>>>

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of discrete randomly occurring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$ when
$N$ have satisfied $b$.
Interpreted with the Binomial distribution of
Equation~\ref{eqn:searches_binomial}, this definition converges to our more
general probability whenever it applies.

We do think of many processes as random when considering scientific data, so
this definition works widely for likelihood functions.
Non-data hypotheses, however, are rarely considered as actually random events,
so the frequency definition cannot not apply.
Prior predictions often apply probabilities to express a subjective uncertainty
predictions that are manifestly non-random.
Under this frequency definition, those predictive probabilities are nonsense
and must be rejected.
The frequency definition does not change probability algebra, it only restricts
it to a permitted scope.

Coin tosses, for example, can be imagined as random.
Earth's human population could toss a few billion coins to precisely
establish a probability of landing heads from the ratio of
$n$ (number of heads) to $N$ (number of tosses).
But suppose you are given a red envelope containing a coin ---
which way up is the coin in the envelope?
You could guess or bet on the answer, and that decision might involve a tacitly
assumed probability.
A frequency, however, would only be permitted if the envelope had been packaged
randomly.
Was the coin first tossed to randomize its face, or placed deliberately?
It is unclear.

Tom Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
stoppard1967rosencrantz
}
contains a clearer example ---
Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads.
They play a game with a third player, landing eight more heads.
Is the next toss heads or tails?
The answer is not random --- it is burned in print onto paper, so no frequency
can be assigned.
Were subjective probabilities permitted, you could predict the result, and that
prediction would be subjectively informed by it being the one-hundred-and-first
toss, our cultural entanglement with in base-10 numbers, and your reading of
these words.
Frequentist theory forbids probabilities of such non-random unknowns, and is
left seeking alternatives.

@@

\TODO{
on ``likelihood'' models ---
we are instructed to pretend that these are likelihoods on auxiliary data
TODO cite.
One is free to engage in such doublethink~\cite{orwell1949nnineteen},
% just as one is free to proclaim that $2 + 2 = 5$
but the fact remains that many constraints do not correspond to predictions of
auxiliary data so are not likelihoods: examples include
scale variations,
theoretical cross-section uncertainties,
non-closure uncertainties (which cover differences between estimates).
simulated sample comparisons,
statistical errors on weighted simulations,
extrapolation uncertainties,
and
all uncertainties on supersymmetric signal yields
(for which we have yet no data).
energy based modelling ---
I know of no tangible benefit from pretending that something is a likelihood
when it is not.
}
